---
output:
  github_document:
    #html_preview: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "README-"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# RAthena

The goal of the RAthena package is to provide a DBI-compliant interface to Athena without the requirement of using odbc and jdbc drivers. This will allow the connection to be athena to be straight forward and only require the installation of python and the AWS SDK boto3.

With the development of AWS sagemaker RAthena will make the connection to AWS Athena simpler.

## Installation:

```{r, eval= F}
remotes::install_bitbucket("shopdirect/rathena",
                           auth_user = "<bitbucket_username>",
                           password = "<bitbucket_password>")
```

## Usage

### Basic Usage

Connect to athena, and send a query and return results back to R.
```{r, eval= F}
library(DBI)

con <- dbConnect(RAthena::athena(),
                aws_access_key_id='YOUR_ACCESS_KEY_ID',
                aws_secret_access_key='YOUR_SECRET_ACCESS_KEY',
                s3_staging_dir='s3://YOUR_S3_BUCKET/path/to/',
                region_name='us-west-2')

res <- dbExecute(con, "SELECT * FROM one_row")
dbFetch(res)
dbClearResult(res)
```

To retrieve query in 1 step.
```{r, eval = F}
dbGetQuery(con, "SELECT * FROM one_row")
```


### Tidyverse Usage
```{r, eval = F}
library(DBI)
library(dplyr)

con <- dbConnect(RAthena::athena(),
                aws_access_key_id='YOUR_ACCESS_KEY_ID',
                aws_secret_access_key='YOUR_SECRET_ACCESS_KEY',
                s3_staging_dir='s3://YOUR_S3_BUCKET/path/to/',
                region_name='us-west-2')
tbl(con, sql("SELECT * FROM one_row"))
```