---
output:
  github_document:
    #html_preview: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "README-"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# RAthena

The goal of the RAthena package is to provide a DBI-compliant interface to Amazon's Athena (https://aws.amazon.com/athena/) using Boto3 SDK. This allows for an efficient, easy setup connection to Athena using the Boto3 SDK as a driver. 

**NOTE:** *Before using RAthena you must have an aws account or have access to aws account with permissions allowing you to use Athena.*

## Installation:

Before installing RAthena ensure that Python 3+ is installed onto your machine: https://www.python.org/downloads/. To install Boto3 either it can installed the pip command or using RAthena installation function:

```
pip install boto3
```
```{R, eval = F}
RAthena::install_boto()
```

To install RAthena (currently not on cran):

```{r, eval= F}
# The development version from Github
remotes::install_github("dyfanjones/rathena")
```


## Usage

### Basic Usage

Connect to athena, and send a query and return results back to R.
```{r, eval= F}
library(DBI)

con <- dbConnect(RAthena::athena(),
                aws_access_key_id='YOUR_ACCESS_KEY_ID',
                aws_secret_access_key='YOUR_SECRET_ACCESS_KEY',
                s3_staging_dir='s3://YOUR_S3_BUCKET/path/to/',
                region_name='eu-west-1')

res <- dbExecute(con, "SELECT * FROM one_row")
dbFetch(res)
dbClearResult(res)
```

To retrieve query in 1 step.
```{r, eval = F}
dbGetQuery(con, "SELECT * FROM one_row")
```

### Intermediate Usage

To create a tables in athena, `dbExecute` will send the query to athena and wait until query has been executed. This makes it and idea method to create tables within athena.

```{r, eval= F}
query <- 
  "CREATE EXTERNAL TABLE impressions (
      requestBeginTime string,
      adId string,
      impressionId string,
      referrer string,
      userAgent string,
      userCookie string,
      ip string,
      number string,
      processId string,
      browserCookie string,
      requestEndTime string,
      timers struct<modelLookup:string, requestTime:string>,
      threadId string,
      hostname string,
      sessionId string)
  PARTITIONED BY (dt string)
  ROW FORMAT  serde 'org.apache.hive.hcatalog.data.JsonSerDe'
      with serdeproperties ( 'paths'='requestBeginTime, adId, impressionId, referrer, userAgent, userCookie, ip' )
  LOCATION 's3://elasticmapreduce/samples/hive-ads/tables/impressions/' ;"
  
dbExecute(con, query)
```

RAthena has 2 extra function to return extra information around Athena tables: `dbGetParitiions` and `dbShow`

`dbGetPartitions` will return all the partitions (returns data.frame):
```{r, eval = F}
RAthena::dbGetPartition(con, "impressions")
```

`dbShow` will return the table's ddl, so you will able to see how the table was constructed in Athena (returns SQL character):
```{r, eval = F}
RAthena::dbShow(con, "impressions")
```

### Advanced Usage

#### Setting up AWS CLI
RAthena is compatible with AWS CLI. This allows your aws credentials to be stored and not be hard coded in your connection.

To install AWS CLI please refer to: https://docs.aws.amazon.com/cli/latest/userguide/cli-chap-install.html, to configure AWS CLI please refer to: https://docs.aws.amazon.com/cli/latest/userguide/cli-chap-configure.html

Once AWS CLI has been set up you will be able to connect to Athena by only putting the `s3_staging_dir`.
```{r, eval = F}
library(RAthena)
con <- dbConnect(RAthena::athena(),
                 s3_staging_dir = 's3://YOUR_S3_BUCKET/path/to/')
```

#### Sending data to Athena
RAthena has created a method to send data.frame from R to Athena.

```{r, eval = F}
# Check existing tables
dbListTables(con)
# Upload iris to Athena
dbWriteTable(con, "iris", iris, 
             partition=c("TIMESTAMP" = format(Sys.Date(), "%Y%m%d")),
             s3.location = "s3://mybucket/data/")

# Read in iris from Athena
dbReadTable(con, "iris")

# Check new existing tables in Athena
dbListTables(con)

# Check if iris exists in Athena
dbExistsTable(con, "iris")
````

### Tidyverse Usage

Creating a connection to Athena and query and already existing table `iris` that was created in previous example.
```{r, eval = F}
library(DBI)
library(dplyr)

con <- dbConnect(RAthena::athena(),
                aws_access_key_id='YOUR_ACCESS_KEY_ID',
                aws_secret_access_key='YOUR_SECRET_ACCESS_KEY',
                s3_staging_dir='s3://YOUR_S3_BUCKET/path/to/',
                region_name='eu-west-1')
tbl(con, sql("SELECT * FROM iris"))
```
```
# Source:   SQL [?? x 5]
# Database: Athena 1.9.210 [eu-west-1/default]
   sepal_length sepal_width petal_length petal_width species
          <dbl>       <dbl>        <dbl>       <dbl> <chr>  
 1          5.1         3.5          1.4         0.2 setosa 
 2          4.9         3            1.4         0.2 setosa 
 3          4.7         3.2          1.3         0.2 setosa 
 4          4.6         3.1          1.5         0.2 setosa 
 5          5           3.6          1.4         0.2 setosa 
 6          5.4         3.9          1.7         0.4 setosa 
 7          4.6         3.4          1.4         0.3 setosa 
 8          5           3.4          1.5         0.2 setosa 
 9          4.4         2.9          1.4         0.2 setosa 
10          4.9         3.1          1.5         0.1 setosa 
# … with more rows
```

dplyr provides lazy querying with allows to short hand `tbl(con, sql("SELECT * FROM iris"))` to `tbl(con, "iris")`. For more information please look at https://db.rstudio.com/dplyr/.
```{r, eval = F}
tbl(con, "iris")
```
```
# Source:   table<iris> [?? x 5]
# Database: Athena 1.9.210 [eu-west-1/default]
   sepal_length sepal_width petal_length petal_width species
          <dbl>       <dbl>        <dbl>       <dbl> <chr>  
 1          5.1         3.5          1.4         0.2 setosa 
 2          4.9         3            1.4         0.2 setosa 
 3          4.7         3.2          1.3         0.2 setosa 
 4          4.6         3.1          1.5         0.2 setosa 
 5          5           3.6          1.4         0.2 setosa 
 6          5.4         3.9          1.7         0.4 setosa 
 7          4.6         3.4          1.4         0.3 setosa 
 8          5           3.4          1.5         0.2 setosa 
 9          4.4         2.9          1.4         0.2 setosa 
10          4.9         3.1          1.5         0.1 setosa 
# … with more rows
```

Querying Athena with `profile_name` instead of hard coding `aws_access_key_id` and `aws_secret_access_key`. By using `profile_name` extra Meta Data is returned in the query to give users extra information.
```{r, eval = F}
con <- dbConnect(RAthena::athena(),
                profile_name = "your_profile",
                s3_staging_dir='s3://YOUR_S3_BUCKET/path/to/')
tbl(con, "iris")) %>% 
  filter(petal_length < 1.3)
```
```
# Source:   lazy query [?? x 5]
# Database: Athena 1.9.210 [your_profile@eu-west-1/default]
   sepal_length sepal_width petal_length petal_width species
          <dbl>       <dbl>        <dbl>       <dbl> <chr>  
 1          4.7         3.2          1.3         0.2 setosa 
 2          4.3         3            1.1         0.1 setosa 
 3          5.8         4            1.2         0.2 setosa 
 4          5.4         3.9          1.3         0.4 setosa 
 5          4.6         3.6          1           0.2 setosa 
 6          5           3.2          1.2         0.2 setosa 
 7          5.5         3.5          1.3         0.2 setosa 
 8          4.4         3            1.3         0.2 setosa 
 9          5           3.5          1.3         0.3 setosa 
10          4.5         2.3          1.3         0.3 setosa 
# … with more rows
```
```{r, eval = F}
tbl(con, "iris") %>% 
  select(contains("sepal"), contains("petal"))
```
```
# Source:   lazy query [?? x 4]
# Database: Athena 1.9.210 [your_profile@eu-west-1/default]
   sepal_length sepal_width petal_length petal_width
          <dbl>       <dbl>        <dbl>       <dbl>
 1          5.1         3.5          1.4         0.2
 2          4.9         3            1.4         0.2
 3          4.7         3.2          1.3         0.2
 4          4.6         3.1          1.5         0.2
 5          5           3.6          1.4         0.2
 6          5.4         3.9          1.7         0.4
 7          4.6         3.4          1.4         0.3
 8          5           3.4          1.5         0.2
 9          4.4         2.9          1.4         0.2
10          4.9         3.1          1.5         0.1
# … with more rows
```

# Similar Projects

## Python:
* `pyAthena` - A python wrapper of the python package `Boto3` using the sqlAlchemy framework: https://github.com/laughingman7743/PyAthena
* `pyAthenaJDBC` - A python interface into AWS Athena's JDBC drivers: https://github.com/laughingman7743/PyAthenaJDBC

## R:
* `AWR.Athena` - A R wrapper of RJDBC for the AWS Athena's JDBC drivers: https://github.com/nfultz/AWR.Athena

## Comparison:
The reason why `RAthena` stands slightly apart from `AWR.Athena` is that `AWR.Athena` uses the Athena JDBC drivers and `RAthena` uses the Python AWS SDK `Boto3`. The ultimate goal is to provide an extra method for R users to interface with AWS Athena. As `pyAthena` is the most similar project, this project has used an appropriate name to reflect this ... `RAthena`.

# To Do list:

* Upload package to cran
* Add a logo (as everyone loves a logo)
* Improve dbFetch method
  * Implement `next_token` in `get_result` for more looping methods.
  * Possibly retrieve metadata from `get_result` for variable conversion (when `n != -1, Inf`)
* Streamline existing functions:
  * Increase current performance (performance testing will need to be added)